\section{Preliminaries}\label{sec:prelim}

This section recalls the threshold-automata framework of Konnov et
al.~\cite{KonnovVW2015, KonnovLVW2017} and establishes notation used
throughout the paper.

\subsection{Threshold Automata}

A \emph{threshold automaton} is a tuple
$\ta = (\mathcal{L}, \mathcal{R}, \Gamma, \Pi, \Phi, \mathcal{I})$
where:
\begin{itemize}
  \item $\mathcal{L}$ is a finite set of \emph{locations} (control
    states of a process template);
  \item $\Pi = \{p_0, \ldots, p_m\}$ is a set of \emph{parameters}
    (e.g.\ $n$, $t$, $f$);
  \item $\Phi$ is a \emph{resilience condition}---a conjunction of
    linear inequalities over~$\Pi$ (e.g.\ $n > 3t \wedge f \leq t$);
  \item $\Gamma = \{\gamma_0, \ldots, \gamma_s\}$ is a set of
    \emph{shared variables} (message counters);
  \item $\mathcal{R}$ is a set of \emph{rules}, each of the form
    $r = (\ell, \ell', g, u)$ with source location~$\ell$, target
    location~$\ell'$, threshold guard~$g$, and update~$u$;
  \item $\mathcal{I} \subseteq \mathcal{L}$ is the set of initial
    locations.
\end{itemize}

A \emph{threshold guard}~$g$ is a conjunction of atoms
$\sum_{i} \gamma_i \geq c_0 + \sum_j c_j \cdot p_j$ where
$c_0, c_j$ are integer constants.  An \emph{update}~$u$ increments
zero or more shared variables by one.

Intuitively, every process executes the same template: it sits in some
location, waits until a threshold guard is satisfied by the global
message counters, then fires a rule that moves it to a new location
and increments the appropriate counters.  The key insight is that
the protocol's correctness depends on \emph{how many} processes are
in each location, not on \emph{which} processes.

\subsection{Counter Abstraction}\label{sec:counter-abstraction}

This symmetry observation motivates \emph{counter
abstraction}~\cite{EmersonNamjoshi1995}: instead of tracking each of
the $n$~processes individually, we track the \emph{number} of processes
in each location.

A \emph{configuration} is a pair $(\vec\kappa, \vec\gamma)$ where
$\kappa[\ell] \in \mathbb{N}$ counts the processes at location~$\ell$
(subject to $\sum_{\ell} \kappa[\ell] = n$) and $\vec\gamma$ records
global message counts.

\paragraph{Step relation.}
At each step~$k$, some number $\delta[k,r] \geq 0$ of processes fire
each rule~$r$.  Location counters update as:
\begin{align}
  \kappa'[\ell] &= \kappa[\ell]
    - \sum_{\{r : \mathit{src}(r)=\ell\}} \delta[k,r]
    + \sum_{\{r : \mathit{dst}(r)=\ell\}} \delta[k,r]
  \label{eq:kappa-update}
\end{align}
Shared variables update according to each rule's update function, plus
an \emph{adversary injection} term $\adv[k,\gamma] \geq 0$ that models
Byzantine message forgery, bounded globally by the fault parameter:
$\sum_{k} \adv[k,\gamma] \leq f$ for each~$\gamma$.

\paragraph{Why this works.}
Konnov et al.~\cite{KonnovLVW2017} showed that for symmetric
threshold-guarded protocols there exists a \emph{cutoff}: a system size
beyond which no new behaviors arise.  Counter abstraction is therefore
\emph{exact} for this class---the symbolic encoding over parameters
$n, t, f$ captures all behaviors of every concrete instantiation
satisfying~$\Phi$.

\subsection{Verification Strategies}

\tarsier{} employs three complementary strategies:

\paragraph{Bounded model checking (BMC).}
Unrolls the step relation for $k$~steps and checks whether a bad state
is reachable~\cite{BiereCCSZ1999}.  Sound for finding bugs; incomplete
for proving their absence.

\paragraph{$k$-Induction.}
Combines BMC with an inductive step: if no bad state is reachable in
$k$~steps from \emph{any} configuration satisfying the safety property,
then the property holds universally~\cite{SheehySSS2000}.

\paragraph{Property-directed reachability (PDR/IC3).}
Iteratively constructs an inductive invariant by blocking predecessors
of bad states~\cite{Bradley2011}.  Fully automatic, often more efficient
than $k$-induction on protocols with complex invariant structure.
