\section{Faithful Network Semantics}\label{sec:semantics}

The central technical contribution of \tarsier{} is a hierarchy of four
\emph{network abstraction modes} that progressively refine the classical
counter abstraction with more faithful message-delivery semantics.  Each
mode trades analysis cost for modeling precision, and we provide formal
soundness guarantees for each.

\subsection{The Faithful Target Model}

We define a \emph{faithful protocol model} as the idealized semantics
where:
\begin{itemize}
  \item every process has a unique identity
    ($\mathit{pid} \in [0, n{-}1]$);
  \item messages are delivered per-recipient (sender $s$ sends to
    recipient~$r$);
  \item channels are sender-authenticated (\texttt{auth: signed});
  \item threshold guards count \emph{distinct senders}, not message
    copies.
\end{itemize}
This is the model that protocol designers have in mind when they write
pseudo-code.  All four abstraction modes approximate it from above (more
behaviors) or match it exactly.

\subsection{Mode Hierarchy}

\paragraph{Mode 1: Classic.}
The original counter abstraction of Konnov et
al.~\cite{KonnovVW2015}.  Shared variables are scoped per (message
family, recipient role, field valuation).  All senders are aggregated
into a single counter; the adversary injects messages globally.  This
is the coarsest---and fastest---abstraction.

\paragraph{Mode 2: Identity-selective.}
Shared variables are refined to per-sender, per-recipient scope.  Each
Byzantine sender receives a \emph{budget variable}
$\mathit{byzsender}[s] \in \{0,1\}$ with
$\sum_s \mathit{byzsender}[s] \leq f$.  Equivocation control
(\texttt{equivocation: none}) constrains each sender to a single
payload per round.  This mode correctly models the distinct-sender
counting that most BFT protocols rely on.

\paragraph{Mode 3: Cohort-selective.}
Recipients are partitioned into \emph{cohorts}---sub-groups that share
delivery constraints.  This provides intermediate granularity between
identity-selective (per-sender) and classic (per-role), useful for
protocols with committee structures.

\paragraph{Mode 4: Process-selective.}
Every process has a concrete identity $\mathit{pid} \in [0, n{-}1]$,
and channels are fully per-process.  The adversary activates a
\emph{static faulty set} $F \subseteq [0, n{-}1]$ with $|F| \leq f$;
only processes in~$F$ may behave Byzantine.  This is the most precise
mode.

\subsection{Soundness Guarantees}\label{sec:transfer}

Let $\mathcal{R}_M(\ta)$ denote the reachable configurations of
automaton~$\ta$ under mode~$M$, and $\mathcal{R}_{\mathit{faith}}$ the
reachable configurations in the faithful target.

\begin{theorem}[Classic soundness]\label{thm:classic}
If guards are monotone ($\geq$, $>$ only) and equivocation is
\texttt{full}, then
$\mathcal{R}_{\mathit{faith}} \subseteq
 \mathcal{R}_{\mathit{classic}}(\ta)$.
A SAFE verdict under classic mode implies safety in the faithful model.
\end{theorem}

This is the standard over-approximation guarantee.  Critically, an
UNSAFE verdict under classic mode does \emph{not} transfer---the
counterexample may be spurious, arising from behaviors that the
faithful model cannot actually produce.

\begin{theorem}[Identity-selective soundness]\label{thm:idsel}
With \texttt{auth: signed} and sender budgets at least as large as
the faithful target's,
$\mathcal{R}_{\mathit{faith}} \subseteq
 \mathcal{R}_{\mathit{id\text{-}sel}}(\ta)$.
\end{theorem}

Identity-selective mode is a \emph{tighter} over-approximation that
correctly captures distinct-sender thresholds and equivocation
constraints.  SAFE verdicts transfer; UNSAFE verdicts still may not.

\begin{theorem}[Process-selective exactness]\label{thm:procsel}
For bounded $\mathit{pid} \in [0, n{-}1]$ with \texttt{auth: signed}:
$\mathcal{R}_{\mathit{proc\text{-}sel}}(\ta) =
 \mathcal{R}_{\mathit{faith}}$.
Both SAFE and UNSAFE verdicts transfer.
\end{theorem}

Process-selective mode eliminates the abstraction gap entirely.  An
UNSAFE verdict corresponds to a \emph{real attack}, making this mode
ideal for precise bug-finding.

\subsection{Adversary Encoding}

For each mode, the adversary's power is encoded in the SMT formula via:
\begin{enumerate}
  \item \textbf{Injection variables:}
    $\adv[k, \gamma] \geq 0$ at each step~$k$, bounded by
    $\sum_k \adv[k, \gamma] \leq f$.
  \item \textbf{Non-forgeability:}
    $\adv[k, \gamma_{\mathit{crypto}}] = 0$ for cryptographic counters.
  \item \textbf{Sender budgets} (identity-selective and above):
    $\mathit{byzsender}[s] \in \{0,1\}$,
    $\sum_s \mathit{byzsender}[s] \leq f$.
  \item \textbf{Static faulty set} (process-selective):
    $\mathit{faulty}[\mathit{pid}] \in \{0,1\}$,
    $\sum_{\mathit{pid}} \mathit{faulty}[\mathit{pid}] \leq f$.
\end{enumerate}

\subsection{Fallback Lattice}

When the SMT formula for a precise mode exceeds solver capacity,
\tarsier{} supports automatic fallback along a lattice:
\[
  \text{process-selective}
  \rightarrow \text{cohort-selective}
  \rightarrow \text{identity-selective}
  \rightarrow \text{classic}
\]
Each step widens the abstraction (admitting more behaviors) while
preserving SAFE-verdict soundness at the cost of UNSAFE-verdict
precision.  The tool reports which mode produced the final verdict so
users can assess the result's strength.
